---
title: "R-club-June28"
author: "Michelle Tang"
date: "6/28/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(nycflights13)
```


### 13.3.1 Exercises

1. Add a surrogate key to flights.

2. Identify the keys in the following datasets

Lahman::Batting,
babynames::babynames
nasaweather::atmos
fueleconomy::vehicles
ggplot2::diamonds
(You might need to install some packages and read some documentation.)

3. Draw a diagram illustrating the connections between the Batting, Master, and Salaries tables in the Lahman package. Draw another diagram that shows the relationship between Master, Managers, AwardsManagers.

How would you characterise the relationship between the Batting, Pitching, and Fielding tables?

### 13.4.6 Exercises

1. Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:

airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
(Don’t worry if you don’t understand what semi_join() does — you’ll learn about it next.)

You might want to use the size or colour of the points to display the average delay for each airport.

2. Add the location of the origin and destination (i.e. the lat and lon) to flights.

3. Is there a relationship between the age of a plane and its delays?

4. What weather conditions make it more likely to see a delay?

5. What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.

### 13.5.1 Exercises

1. What does it mean for a flight to have a missing tailnum? What do the tail numbers that don’t have a matching record in planes have in common? (Hint: one variable explains ~90% of the problems.)

2. Filter flights to only show flights with planes that have flown at least 100 flights.

3. Combine fueleconomy::vehicles and fueleconomy::common to find only the records for the most common models.

4. Find the 48 hours (over the course of the whole year) that have the worst delays. Cross-reference it with the weather data. Can you see any patterns?

5. What does anti_join(flights, airports, by = c("dest" = "faa")) tell you? What does anti_join(airports, flights, by = c("faa" = "dest")) tell you?

6. You might expect that there’s an implicit relationship between plane and airline, because each plane is flown by a single airline. Confirm or reject this hypothesis using the tools you’ve learned above.

