---
title: "R4ds_May10"
author: "Michelle Tang"
date: "5/9/2017"
output: html_document
---

##Data transformation

```{r setup, include=FALSE}
library(tidyverse)
library(nycflights13)
```

```{r}
flights
```

**Five Key Functions**

*Pick observations by their values (filter()).
*Reorder the rows (arrange()).
*Pick variables by their names (select()).
*Create new variables with functions of existing variables (mutate()).
*Collapse many values down to a single summary (summarise()).

```{r}
filter(flights, month==1, sched_dep_time==515)
```

For other types of combinations, you’ll need to use Boolean operators yourself: & is “and”, | is “or”, and ! is “not”.

== versus near()
near() is an approximation

###5.2.4 Exercises

1. Find all flights that

  + Had an arrival delay of two or more hours
  
```{r}
filter(flights, arr_delay >=120)
```

  + Flew to Houston (IAH or HOU)
  
```{r}
filter(flights, dest=="IAH" | dest=="HOU")
```

  + Were operated by United, American, or Delta

```{r}
filter(flights, carrier == "UA" | carrier == "AA" | carrier =="DL")
```

  + Departed in summer (July, August, and September)
  
```{r}
filter(flights, month %in% c(7,8,9))
```

  + Arrived more than two hours late, but didn’t leave late
  
```{r}
filter(flights, arr_delay >= 120 & dep_delay <= 0)
```

  + Were delayed by at least an hour, but made up over 30 minutes in flight (not sure what the second half means)
  
```{r}
filter(flights, dep_delay >= 60 & sched_arr_time-arr_time > 30)
```

  
  + Departed between midnight and 6am (inclusive)
  
```{r}
filter(flights, dep_time >= 0 & dep_time <= 600)
```


2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

includes values inside upper and lower boundaries.

```{r}
?between
filter(flights, between(dep_time, 0, 600))
```

3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
filter(flights, is.na(dep_time))
```

Other variables that are missing data is dep_delay, arr_time, arr_delay, air_time
These might be canceled flights

4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

Because anything to the 0 power is 1
Because NA | TRUE includes TRUE which is not missing
FALSE, like NA is a logical object. 
Input of logical object and NA with logical operators will return another logical object
NA * 0 might not work because NA could be anything, not just a number and not a logical object

###5.3.1 Exercises

1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).

```{r}
arrange(flights, desc(is.na(dep_time)))
```


2. Sort flights to find the most delayed flights. Find the flights that left earliest.

```{r}
arrange(flights, desc(dep_delay), dep_time)
```

3. Sort flights to find the fastest flights.

```{r}
arrange(flights, air_time)
```

4. Which flights travelled the longest? Which travelled the shortest?

```{r}
arrange(flights, desc(distance))
arrange(flights, distance)
```

###5.4.1 Exercises

1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.

```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
select(flights, dep_time, dep_delay:arr_time, arr_delay)
select(flights, -(year:day), -(sched_dep_time),-(sched_arr_time), -(carrier:time_hour))
```

2. What happens if you include the name of a variable multiple times in a select() call?

```{r}
select(flights, day, day, day)
select(flights, carrier, day, carrier)
```

It gets selected only once.

3. What does the one_of() function do? Why might it be helpful in conjunction with this vector?

```{r}
?one_of
```

Allows selecting variables contained in a character vector.

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")

select(flights, one_of(vars))
```


4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
select(flights, contains("TIME"))
```

By default, contains() ignores case in argument ignore.case = TRUE. Can change to FALSE if need to match case.

###5.5.2 Exercises

1. Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
transmute(flights, dep_time, dep_midnight = dep_time %/% 100 * 60 + dep_time %% 100)
transmute(flights, sched_dep_time, sched_dep_midnight = sched_dep_time %/% 100 * 60 + sched_dep_time %% 100)
```


2. Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

```{r}
select(flights, air_time) 
mutate(flights, arr_time - dep_time)
```

They don't equal to each other.
To fix it, need to change arr_time and dep_time to minutes after minute and then substract.


3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

```{r}
select(flights, dep_time, sched_dep_time, dep_delay)
```

delay_delay = dep_time - sched_dep_time

4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

```{r, echo=FALSE}
#min_rank(desc(flights$dep_delay)) #this gives you ranking
```

To handle ties, can use variation of min_rank(). min_rank has gaps in ranks for ties. Can use row_number or dense_rank

5. What does 1:3 + 1:10 return? Why?

```{r}
#1:3 + 1:10
```

It cycles 1 through 3 and adds to 1 through 10 but it gives error because 10 isn't a mulitple of 3

6. What trigonometric functions does R provide?

```
cos(x)
sin(x)
tan(x)

acos(x)
asin(x)
atan(x)
atan2(y, x)

cospi(x)
sinpi(x)
tanpi(x)
```
