---
title: "R-club-May17"
author: "Michelle Tang"
date: "5/15/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

For May 17, we'll finish the 'Data transformation' sections (group summaries and group mutates) and the very brief 'Workflow: scripts' chapter.'

### 5.6

Grouped summaries with summarise()

Pair it with group_by() is more useful

```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

Combining multiple operations with pipe

```{r}
#code you might write with dplyr (which is now part of tidyverse)
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)
delay <- filter(delay, count > 20, dest != "HNL")

# It looks like delays increase with distance up to ~750 miles 
# and then decrease. Maybe as flights get longer there's more 
# ability to make up delays in the air?
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)

```

```{r}
#example using pipe "%>%"

delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
```

```{r}
#Are they the same?
delays == delay

#Yes
```

Missing values

Use na.rm = True to remove missing values before computation

Available wtih all aggregation functions

Get a count of non-missing values sum(!(is.na(x)))

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))

delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay)
  )

ggplot(data = delays, mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)
```

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )

ggplot(data = delays, mapping = aes(x = n, y = delay)) + 
  geom_point(alpha = 1/10)
```

```{r}
delays %>% 
  filter(n > 25) %>% 
  ggplot(mapping = aes(x = n, y = delay)) + 
    geom_point(alpha = 1/10)
```

```{r}
#install.packages("Lahman")
#library(Lahman)
batting <- as_tibble(Lahman::Batting)

batters <- batting %>% 
  group_by(playerID) %>% 
  summarise(
    ba = sum(H, na.rm = TRUE) / sum(AB, na.rm = TRUE),
    ab = sum(AB, na.rm = TRUE)
  )

batters %>% 
  filter(ab > 100) %>% 
  ggplot(mapping = aes(x = ab, y = ba)) +
    geom_point() + 
    geom_smooth(se = FALSE)

batters %>% arrange(desc(ba))
```

Useful summary functions

**Measure of spread: sd(x), IQR(x), mad(x)**

IQR() interquartile range

mad(x) median absolute deviation

**Measure of rank: min(x), quantile(x, 0.25), max(x)**

quantile(x, 0.25) find values of x that is greater than 25%5 of the values and less thn the remaining 75%

**Measures of position: first(x), nth(x,2), last(x)**

similar to subset, e.g. x[1], x[length(x)]

```{r}
daily <- group_by(flights, year, month, day)
(per_day   <- summarise(daily, flights = n()))
per_month <- summarise(per_day, flights = sum(flights))
per_year  <- summarise(per_month, flights = sum(flights))
```


### 5.6.7 Exercises

1. Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:

A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.

```{r}
flights %>% group_by(flight) %>% summarise(arrival=median(arr_delay)) %>% filter(arrival==-15)

flights %>% group_by(flight) %>% summarise(arrival=median(arr_delay)) %>% filter(arrival==15)
```


A flight is always 10 minutes late.

A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.

99% of the time a flight is on time. 1% of the time it’s 2 hours late.

Which is more important: arrival delay or departure delay?

2. Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).

3. Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?

4. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?

5. Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))

6. What does the sort argument to count() do. When might you use it?
