---
title: "R-club-May17"
author: "Michelle Tang"
date: "5/15/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

For May 17, we'll finish the 'Data transformation' sections (group summaries and group mutates) and the very brief 'Workflow: scripts' chapter.'

### 5.6

Grouped summaries with summarise()

Pair it with group_by() is more useful

```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

Combining multiple operations with pipe

```{r}
#code you might write with dplyr (which is now part of tidyverse)
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)
delay <- filter(delay, count > 20, dest != "HNL")

# It looks like delays increase with distance up to ~750 miles 
# and then decrease. Maybe as flights get longer there's more 
# ability to make up delays in the air?
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)

```

```{r}
#example using pipe "%>%"

delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
```

```{r}
#Are they the same?
delays == delay

#Yes
```

Missing values

Use na.rm = True to remove missing values before computation

Available wtih all aggregation functions

Get a count of non-missing values sum(!(is.na(x)))

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))

delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay)
  )

ggplot(data = delays, mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)
```

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )

ggplot(data = delays, mapping = aes(x = n, y = delay)) + 
  geom_point(alpha = 1/10)
```

```{r}
delays %>% 
  filter(n > 25) %>% 
  ggplot(mapping = aes(x = n, y = delay)) + 
    geom_point(alpha = 1/10)
```

```{r}
#install.packages("Lahman")
#library(Lahman)
batting <- as_tibble(Lahman::Batting)

batters <- batting %>% 
  group_by(playerID) %>% 
  summarise(
    ba = sum(H, na.rm = TRUE) / sum(AB, na.rm = TRUE),
    ab = sum(AB, na.rm = TRUE)
  )

batters %>% 
  filter(ab > 100) %>% 
  ggplot(mapping = aes(x = ab, y = ba)) +
    geom_point() + 
    geom_smooth(se = FALSE)

batters %>% arrange(desc(ba))
```

Useful summary functions

**Measure of spread: sd(x), IQR(x), mad(x)**

IQR() interquartile range

mad(x) median absolute deviation

**Measure of rank: min(x), quantile(x, 0.25), max(x)**

quantile(x, 0.25) find values of x that is greater than 25%5 of the values and less thn the remaining 75%

**Measures of position: first(x), nth(x,2), last(x)**

similar to subset, e.g. x[1], x[length(x)]

```{r}
daily <- group_by(flights, year, month, day)
(per_day   <- summarise(daily, flights = n()))
per_month <- summarise(per_day, flights = sum(flights))
per_year  <- summarise(per_month, flights = sum(flights))
```


### 5.6.7 Exercises

1. Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:

A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.

```{r}
flights %>% group_by(flight) %>% summarise(early = mean(arr_delay == -15)) %>% filter(early==1)

flights %>% group_by(flight) %>% summarise(late = mean(arr_delay == 15)) %>% filter(late == 1)
```


A flight is always 10 minutes late.

```{r}
flights %>% group_by(flight) %>% summarise(late = mean(arr_delay==10)) %>% filter(late ==1)
```


A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.

Same as first example

```{r}
flights %>% group_by(flight) %>% summarise(early=mean(arr_delay==-30)) %>% filter(early==1)

flights %>% group_by(flight) %>% summarise(late=mean(arr_delay==30)) %>% filter(late==1)
```


99% of the time a flight is on time. 1% of the time it’s 2 hours late.

```{r}

```


Which is more important: arrival delay or departure delay?

Mean of delay time, median of delay time, spread (sd) of delay time, how flights rank in terms of delay time, .

2. Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).

```{r}
not_cancelled %>% count(dest)

not_cancelled %>% group_by(dest) %>% summarise(n=n())

not_cancelled %>% count(tailnum, wt = distance)

not_cancelled %>% group_by(tailnum) %>% summarise(wt = sum(distance))
```


3. Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?

Longer to type. Could easily indicate cancelled flights as is.na(air_time)

4. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?

```{r}
cancel<-flights %>% group_by(month, day) %>% count()

ggplot(cancel, mapping = aes(day, n)) + geom_point()

canceldelay <- flights %>% mutate(cancelled = (is.na(air_time))) %>% group_by(month, day) %>% summarise(cancel = mean(cancelled), avg_delay = mean(dep_delay, na.rm = T))

ggplot(canceldelay, aes(avg_delay, cancel)) + geom_point() + geom_smooth()
```

Weak positive relationship

5. Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))

```{r}
flights %>% group_by(carrier, dest) %>% summarise(n())

flights %>% group_by(carrier, dest) %>% summarise(dep=mean(dep_delay), arr=mean(arr_delay)) %>% arrange(desc(arr))

delay_carr<-flights %>%
  group_by(carrier) %>%
  summarise(dep_delay=mean(dep_delay, na.rm=T),arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(arr_delay))

delay_dest<-flights %>%
  group_by(dest) %>%
  summarise(dep_delay=mean(dep_delay, na.rm=T), arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(arr_delay))
```

F9 carrier has the worst delays

Can disentangle bad carrier versus bad airports by comparing avg delay by carrier versus avg delay by dest. Those averages also don't correspond to delay calculated by carrier and dest


6. What does the sort argument to count() do. When might you use it?

```{r}
?count
```

Sort, if true, will sort output in descending order of n. Might want to use it find max.

```{r}
flights %>% count(dest, sort=T)
```



### 5.7.1 Exercises

1. Refer back to the lists of useful mutate and filtering functions. Describe how each operation changes when you combine it with grouping.

Would apply to your defined group instead of over the entire df/tibble.

2. Which plane (tailnum) has the worst on-time record?

```{r}
flights %>% group_by(tailnum) %>% summarise(arr_delay = mean(arr_delay, na.rm=T)) %>% arrange(desc(arr_delay))
```

N844MH worst on-time record.

3. What time of day should you fly if you want to avoid delays as much as possible?

```{r}
flights %>% group_by(hour) %>% summarise(arr_delay = mean(arr_delay, na.rm=T)) %>% arrange(arr_delay)
```

Around 7 am.

4. For each destination, compute the total minutes of delay. For each, flight, compute the proportion of the total delay for its destination.

```{r}
flights %>% filter(!(is.na(arr_delay))) %>% group_by(dest) %>% summarise(total = sum(arr_delay))

#need to group by carrier and flight
flights %>% filter(!(is.na(arr_delay))) %>% group_by(dest) %>% mutate(proportion = arr_delay/sum(arr_delay))

```

5. Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using lag() explore how the delay of a flight is related to the delay of the immediately preceding flight.

```{r}
lag_flights<-flights %>% group_by(month, day, hour) %>% filter(!is.na(arr_delay)) %>% mutate(lag_d=lag(dep_delay))

ggplot(lag_flights, aes(dep_delay, lag_d)) + geom_point()
```


6. Look at each destination. Can you find flights that are suspiciously fast? (i.e. flights that represent a potential data entry error). Compute the air time a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?

Some of these questions are very confusing . . . compute air time a flight relative to shortest flight to that dest??

```{r}
flights %>% filter(!(is.na(arr_delay))) %>% group_by(dest) %>% summarise(mean_air=mean(air_time)) %>% arrange(mean_air)
```

Flights to BDL, not surprising if it's in Connecticutt.


```{r}
flights %>% filter(!(is.na(arr_delay))) %>% group_by(dest) %>% mutate(avg_air=mean(air_time), min_air=min(air_time), fast = air_time-min_air) %>% arrange(desc(fast))
```


7. Find all destinations that are flown by at least two carriers. Use that information to rank the carriers.

```{r}
flights %>% group_by(dest) %>% count(carrier) %>% summarise(num_carrier=n()) %>% filter(num_carrier >= 2)
```


8. For each plane, count the number of flights before the first delay of greater than 1 hour.


```{r}
flights %>% arrange(month, day, tailnum) %>% group_by(tailnum) %>% mutate(hourlate = dep_delay > 60) %>% mutate(beforedelay = cumsum(hourlate)) %>% filter(beforedelay < 1) %>% count()
```

### 6.3 

1. Go to the RStudio Tips twitter account, https://twitter.com/rstudiotips and find one tip that looks interesting. Practice using it!

code snippets: https://rviews.rstudio.com/2017/03/03/february-17-tips-and-tricks/

```{r}
#template for r function is fun + Tab
name <- function(variables) {
  
}
```


2. What other common mistakes will RStudio diagnostics report? Read https://support.rstudio.com/hc/en-us/articles/205753617-Code-Diagnostics to find out.

Missing argument

If variable not in scope

If variable is defined but not used