---
title: "R_Club_DataVis"
author: "Michelle Tang"
date: "4/24/2017"
output: html_document
---
Exercises from 'R for Data Science', Data Visualization to end of 3.5.1.

3.1.1

Load tidyverse, install package if needed.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

3.2.1
Do cars with big engines use more fuel than cars with small engines?

```{r}
mpg<-mpg
head(mpg)
```

3.2.2 through 3.2.4

Scatterplot
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x=displ, y = hwy))
```

Graphing template
ggplot(data = <DATA>) + <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))

####3.2.4 Exercises

1. ggplot(data = mpg)

```{r}
ggplot(data = mpg)
```

An empty field

2. How many rows and columns in mtcar

```{r}
dim(mtcars)
```

3. Describe drv variable

```{r}
?mpg
```

f = front wheel drive
r = rear wheel drive
4 = 4wd

4. Scatterplot of hwy vs cyl

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = hwy, y = cyl))
```

5. Scatterplot class vs drv

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x=class, y = drv))
```

Not useful as I don't see any trends or linear relationship.

3.3 Aesthetic mapping
```{r}
ggplot(data = mpg) + geom_point(mapping=aes(x=displ, y=hwy, color = class))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))

```

```{r}
# Left
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```

```{r}
# Right
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

####3.3.1 Exercises
1. What’s gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

The color of points isn't blue. The color = "blue" is inside aes() so tried to map aesthics to a variable called "blue" which doesn't exist.

2. Which variables in mpg are categorical? Which continuous?
```{r}
str(mpg)
```

```{r}
?mpg
```

Categorical: manufacturer, model, class, fl, drv, trans, year, cyl
Continuous: cty, hwy, displ, year

3. Map a continuous variable to color, size and shape. How does these aesthetics behave?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x=displ, y = hwy, color = cty)) #continuous
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x=displ, y = hwy, color = drv)) #categorical
```


```{r}
#ggplot(data = mpg) + geom_point(mapping = aes(x=displ, y = hwy, shape = cty)) #continuous
```
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x=displ, y = hwy, shape = drv)) #categorical
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x=displ, y = hwy, size = cty)) #continuous
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x=displ, y = hwy, size = drv)) 
```

Color: Gradient for continuous, discrete aesthetics for categorical. 
Size: Gradient for continous and categorical
Shape: Error for continuous and discrete for categorical

4. Map same variable to multiple aesthetics

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x=displ, y = hwy, color = cty, size = cty)) 
```
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x=displ, y = hwy, color = drv, shape=drv)) 
```

Possbile. For categorical, multiple aesthetics can be mapped and given one legend. FOr continous, excluding shape, get two legends.

5. What does stroke aesthetic do?
```{r}
?geom_point
```

To modify width of border for shapes that have a border.

```{r}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point(shape = 21, colour = "black", fill = "white", size = 5, stroke = 5)
```
```{r}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point(shape = 21, colour = "black", fill = "white", size = 5)
```

6. What happens if you map aesthic to somethign other than a variable?

```{r}
ggplot(data = mpg) +
  geom_point(mapping= aes(x=trans, y = hwy, color = displ < 5))
```

Becomes a boolean 

3.5 Facets

To split into subplots, categorical variables
facet_wrap(~ <VARIABLE>)

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

Can facet by two variables 
facet_grid(<VARIABLE1> ~ <VARIABLE2>)
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```

####3.5.1 Exercises
1. What happens if facet on a continuous variable?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ hwy)
```
Makes each number in continuous variable into discrete category

2. What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x=drv, y=cyl)) + facet_grid(drv~cyl)
```
Empty cells mean there is not data point for that coordinate of drv, cyl. The empty cells relate to the absence of data point in that coordinate.

3. What plots does the following code make? What does . do?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
## scatterplot of displ versus hwy, faceted based on wheel drive

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
## scatterplot of disp versus hwy, faceted based on # of cyl
```

The . prevents faceting in row or columns dimension when faceting. From ?facet_grid: "the dot in the formula is used to indicate there should be no faceting on this dimension (either row or column)."

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ cyl)
```

but how is it different from below: if you ommit the .
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(~ cyl)
```

4. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

Adv: Less crowded, can look at data separated by subsetting
Dis: Taken away from whole, hard to see how different subsets of data might be driving whole data trends
More faceting as dataset gets larger

5. Read ?facet_wrap
```{r}
?facet_wrap
```

nrow indicate number of rows for facets
ncol indicate number of cols for facets

facet_grid doesn't have nrow or ncol variables because that's pre-defined in the facetting variables for facets 

6. When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

Better use of screen space. Computer screens wider than they are tall.