---
title: "R-Club-May31"
author: "Michelle Tang"
date: "5/30/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


## 7.6 Patterns and models

```{r}
ggplot(data = faithful) + 
  geom_point(mapping = aes(x = eruptions, y = waiting))
```

Fits a model that predicts price from carat and computes the residuals (difference between the predicted value and actual value)

```{r}
library(modelr)

mod <- lm(log(price) ~ log(carat), data = diamonds)

diamonds2 <- diamonds %>% 
  add_residuals(mod) %>% 
  mutate(resid = exp(resid))

ggplot(data = diamonds2) + 
  geom_point(mapping = aes(x = carat, y = resid))
```

See residuals

```{r}
ggplot(data = diamonds2) + 
  geom_boxplot(mapping = aes(x = cut, y = resid))
```

## 7.7 ggplot2 calls

The first two arguments to `ggplot()` are `data` and `mapping`.

The first two arguments to `aes()` are `x` and `y`.

## 8 Workflow

Best to consider your R scripts as the final product because it's easier to recreate the environment.

Cmd + Shift + F10 to retart RStudio

Cmd + Shift + S to rerun current script

Don't use absolute paths because it hinders sharing. To address this problem, create new projects and then you can use relative paths.

## 8.5 Summary

In summary, RStudio projects give you a solid workflow that will serve you well in the future:

* Create an RStudio project for each data analysis project.

* Keep data files there; we’ll talk about loading them into R in data import.

* Keep scripts there; edit them, run them in bits or as a whole.

* Save your outputs (plots and cleaned data) there.

* Only ever use relative paths, not absolute paths.

Everything you need is in one place, and cleanly separated from all the other projects that you are working on.

## Tibbles

Creating tibbles

```{r}
as_tibble(iris)
tibble(
  x = 1:5, 
  y = 1, 
  z = x ^ 2 + y
)
```


Non-syntactic names:

Not starting with a letter or contain unusual characters like a space.

Transposed tibble `tribble()`

Column heading defined by formulas and entries are separated by commas.

```{r}
tribble(
  ~x, ~y, ~z,
  #--|--|----
  "a", 2, 3.6,
  "b", 1, 8.5
)
```

## 10.3 Tibbles vs data.frame

Two main differences in usages of tibble vs classic data.frame: printing and subsetting

Tibbles show only the first 10 rows and all the columns that fit on the screen and each column reports its type.

Can control number of rows and width of display with `print()`

```{r}
nycflights13::flights %>% 
  print(n = 10, width = Inf)
```

Control print behavior:

Print n number of rows
`options(tibble.print_max = n, tibble.print_min = m)`

Always print all columns
`options(tibble.width = Inf)`

Subsetting: `$`, `[[]]`

```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)

# Extract by name
df$x
#> [1] 0.434 0.395 0.548 0.762 0.254
df[["x"]]
#> [1] 0.434 0.395 0.548 0.762 0.254

# Extract by position
df[[1]]
```

Combine with pipe:

```{r}
df %>% .$x
#> [1] 0.434 0.395 0.548 0.762 0.254
df %>% .[["x"]]
#> [1] 0.434 0.395 0.548 0.762 0.254
```

To interaction with older code, turn tibble back to data frame with `as.data.frame()`

## 10.5 Exercises

1. How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).

You can print the object. If it is a tibble, it will only print the first 10 rows and n number of columns that will fit in a screen.

2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

As a data.frame

```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x #this returns the value for xyz even though it's the wrong subset
df[, "xyz"] #returns factor
df[, c("abc", "xyz")] #returns the data frame
```

As a tibble

```{r}
df_t <- tibble(abc = 1, xyz = "a")
df_t$x #gives warning/NULL
df_t[, "xyz"] #returns a tibble
df_t[, c("abc", "xyz")] #returns the tibble but xyz is chr instead of fctr
``` 

3. If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?

```{r}
var <- "xyz"

df_t[[var]]
```
Double brackets

4. Practice referring to non-syntactic names in the following data frame by:

```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
```

Extracting the variable called 1.

```{r}
annoying[['1']]
```


Plotting a scatterplot of 1 vs 2.

```{r}
ggplot(annoying) + geom_point(aes(annoying[['1']], annoying[['2']]))
```


Creating a new column called 3 which is 2 divided by 1.

```{r}
annoying[['3']]<-annoying[['2']]/annoying[['1']]
annoying
```


Renaming the columns to one, two and three.

```{r}
colnames(annoying)<-c("one","two","three")
annoying
```


5. What does tibble::enframe() do? When might you use it?

```{r}
?tibble::enframe
enframe(1:3)
enframe(c(a = 5, b = 7))
```

Converts named atomic vectors or lists to two-column data frames. 

6. What option controls how many additional column names are printed at the footer of a tibble?

`tibble.max_extra_cols` Number of extra columns printed in reduced form.

```{r}
nycflights13::flights
```

```{r}
options(tibble.max_extra_cols = 3)
nycflights13::flights
```

